### 2014-11-10 contract-language.tex
Fixed Typo
Ran spellcheck.

The syntax of our core contract language is shown in Figure
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed effect variables. We use a
special effect variable ($\cureff$) to denote the effect of \emph{current
operation} - the operation for which a contract is being written. The type of
an effect is simply the name of the operation (eg: \cf{withdraw}) that induced
the effect. We admit disjuntion in types to let an effect variable range over
multiple operation names. The contract $\small \forall (a : \tau_1 \vee
\tau_2).~\psi$ is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1}
\vee \oper{a}{\tau_2}) \Rightarrow \psi$.

@@

\item \textbf{Eventually consistency}: Eventually consistent operations can
	be satisfied as long as the client can reach at least one replica. In the
	bank account example, \cf{deposit} is an eventually consistent operation.
	While eventually consistent data stores typically offer \emph{basic} eventual
	consistency with all possible anomalies, we assume that our store provides
	stronger semantics that remain highly-available~\cite{BailisHAT,COPS}; the
	store always exposes a \emph{causal cut} of the updates. This semantics can
	be formally captured in terms of the following contract definition:

@@

The $\Delta$ referred in the above defintion is a conjunction of assumptions
about the nature of primitive relations. A \emph{well-formed} axiomatic
execution ($\E$) is expected to satisfy these assumptions (i.e., $\E \models
\Delta$).

>>>>

The syntax of our core contract language is shown in Figure
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed effect variables. We use a
special effect variable ($\cureff$) to denote the effect of \emph{current
operation} - the operation for which a contract is being written. The type of
an effect is simply the name of the operation (eg: \cf{withdraw}) that induced
the effect. We admit disjunction in types to let an effect variable range over
multiple operation names. The contract $\small \forall (a : \tau_1 \vee
\tau_2).~\psi$ is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1}
\vee \oper{a}{\tau_2}) \Rightarrow \psi$.

@@

\item \textbf{Eventual consistency}: Eventually consistent operations can
	be satisfied as long as the client can reach at least one replica. In the
	bank account example, \cf{deposit} is an eventually consistent operation.
	While eventually consistent data stores typically offer \emph{basic} eventual
	consistency with all possible anomalies, we assume that our store provides
	stronger semantics that remain highly-available~\cite{BailisHAT,COPS}; the
	store always exposes a \emph{causal cut} of the updates. This semantics can
	be formally captured in terms of the following contract definition:

@@

The $\Delta$ referred in the above definition is a conjunction of assumptions
about the nature of primitive relations. A \emph{well-formed} axiomatic
execution ($\E$) is expected to satisfy these assumptions (i.e., $\E \models
\Delta$).

### 2014-11-10 intro.tex
Ran spellcheck.

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency,
availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
modern web services, which aim to provide an "always-on" experience,
overwhelmingly favour availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

@@

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a shallow extension of Haskell that supports the
	description and validation of replicated data types found on eventually
	consistent stores. Contracts are used to specify fine-grained
	application-level consistency properties, and are statically analyzed to
	assign the most efficient and sound store consistency level to the
	corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item An implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.
	Experimental evaluation over a set of real-world applications, including a
	Twitter-like micro-blogging site and an eBay-like auction site illustrates
	the practicality of our approach.
\end{itemize}

The rest of the paper is organized as follows. The next section describes the
system model.  We describe the challenges in programming with eventually
consistent data stores, and introduces \name\ contracts as a proposed solution
to overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
more details the contract language, and its mapping to the store consistency
levels, along with metatheory for certifying the correctness of the mapping.
\S~\ref{sec:txns} introduces transaction contracts and classification.
\S~\ref{sec:impl} describes the implementation and provides details about the
optimizations needed to make the system practical. \S~\ref{sec:results}
discusses experimental evaluation. \S~\ref{sec:related} and~\ref{sec:concl}
present related work and conclusions.

>>>>

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linearizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency,
availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
modern web services, which aim to provide an "always-on" experience,
overwhelmingly favor availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

@@

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a shallow extension of Haskell that supports the
	description and validation of replicated data types found on eventually
	consistent stores. Contracts are used to specify fine-grained
	application-level consistency properties, and are statically analyzed to
	assign the most efficient and sound store consistency level to the
	corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct isolation level for a transaction.
\item We provide meta-theory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item An implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.
	Experimental evaluation over a set of real-world applications, including a
	Twitter-like micro-blogging site and an eBay-like auction site illustrates
	the practicality of our approach.
\end{itemize}

The rest of the paper is organized as follows. The next section describes the
system model.  We describe the challenges in programming with eventually
consistent data stores, and introduces \name\ contracts as a proposed solution
to overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
more details the contract language, and its mapping to the store consistency
levels, along with meta-theory for certifying the correctness of the mapping.
\S~\ref{sec:txns} introduces transaction contracts and classification.
\S~\ref{sec:impl} describes the implementation and provides details about the
optimizations needed to make the system practical. \S~\ref{sec:results}
discusses experimental evaluation. \S~\ref{sec:related} and~\ref{sec:concl}
present related work and conclusions.

### 2014-11-10 motivation.tex
Added basic axioms
Ran spellcheck.

\begin{codehaskell}
type Operation = [e] -> a -> (r, Maybe e)
\end{codehaskell}

\noindent It takes a list of effects (the \emph{context} for the operation),
and an input argument, and returns a result along with an optional effect
(read-only operations return \cf{Nothing}). The new effect (if emitted) is
added to the state of the object at the current replica, and asynchronously
sent to other replicas. The implementation of the bank account operations in
\name is given in Figure~\ref{fig:ex}:

@@

It is easy to see that preventing concurrent \cf{withdraw} operations
eliminates this anamoly. This can be done by insisting that \cf{withdraw} be
executed as a strongly consistent operation. Despite this strengthening,
\cf{getBalance} operation may still incorrectly report a negative balance to
the end user. Consider the execution shown in
fig.~\ref{fig:negativeBalanceAnomaly}, which consists of three concurrent
sessions performing a \cf{deposit}, a \cf{withdraw}, and a \cf{getBalance}
operation, respectively, on the same bank account object. As the \cf{vis} edge
indicates, operation \cf{withdraw(50)} in session 2, witnesses the effects of
\cf{deposit(100)} from session 1, concludes that there is sufficient balance,
and completes successfully. However, the \cf{getBalance} operation may only
witness this successful withdraw, but not the causally preceding \cf{dopsit},
and reports the balance of negative 50 to the user.

@@

In our running example, it is clear that in order to preserve the integrity
constraint, the \cf{withdraw} operation must be strongly consistent.  That is,
given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
vice versa. We express this application-level consistency requirement as a
contract ($\cv_w$) over \cf{withdraw}:

@@

We devise a static approach instead. Contracts are analyzed with the help of a
theorem prover, and statically mapped to a particular store-level consistency
property that the prover guarantees preserves contract semantics. We call this
procedure \emph{contract classification}. Given the variety and complexity of
store level consistency properties, the idea is that the system implementor
parameterizes the classification procedure by describing the store semantics in
the \emph{same} contract language as the one used to express the contract on
the operations. In the next section, we describe the contract language in
detail and describe the classification procedure for a particular store
semantics.

>>>>

\begin{codehaskell}
type Operation = forall e a r. [e] -> a -> (r, Maybe e)
\end{codehaskell}

\noindent It takes a list of effects (the \emph{context} for the operation),
and an input argument, and returns a result along with an optional effect
(read-only operations return \cf{Nothing}). The new effect (if emitted) is
added to the state of the object at the current replica, and asynchronously
sent to other replicas. The implementation of the bank account operations in
\name is given in Figure~\ref{fig:ex}.

@@

It is easy to see that preventing concurrent \cf{withdraw} operations
eliminates this anomaly. This can be done by insisting that \cf{withdraw} be
executed as a strongly consistent operation. Despite this strengthening,
\cf{getBalance} operation may still incorrectly report a negative balance to
the end user. Consider the execution shown in
fig.~\ref{fig:negativeBalanceAnomaly}, which consists of three concurrent
sessions performing a \cf{deposit}, a \cf{withdraw}, and a \cf{getBalance}
operation, respectively, on the same bank account object. As the \cf{vis} edge
indicates, operation \cf{withdraw(50)} in session 2, witnesses the effects of
\cf{deposit(100)} from session 1, concludes that there is sufficient balance,
and completes successfully. However, the \cf{getBalance} operation may only
witness this successful withdraw, but not the causally preceding \cf{deposit},
and reports the balance of negative 50 to the user.

@@

In our running example, it is clear that in order to preserve the integrity
constraint, the \cf{withdraw} operation must be strongly consistent.  That is,
given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
vice-versa. We express this application-level consistency requirement as a
contract ($\cv_w$) over \cf{withdraw}:

@@

We devise a static approach instead. Contracts are analyzed with the help of a
theorem prover, and statically mapped to a particular store-level consistency
property that the prover guarantees preserves contract semantics. We call this
procedure \emph{contract classification}. Given the variety and complexity of
store level consistency properties, the idea is that the system implementer
parameterizes the classification procedure by describing the store semantics in
the \emph{same} contract language as the one used to express the contract on
the operations. In the next section, we describe the contract language in
detail and describe the classification procedure for a particular store
semantics.
